Long term cloud enhancement events of global solar irradiance over Thessaloniki, Greece

### 2024-06-11 article/article.Rmd
Commit 2024-06-11 23:11

Another interesting aspect of the CE events is their distribution within the year.
Figure\nobreakspace{}\ref{fig:relative-month-occurrences} shows the monthly box and
whisker plot of the CE number of occurrence normalized with the highest median value,
that occurs on June.  Although CE events are present  throughout the year, the most
active period  is during May and June.  During the winter (December -- February) the
number of CE cases is about $25\,\%$ of the maximum.  For the rest months, the
occurrences gradually ramp between the maximum and minimum.  This seasonality is a
combinations of the clouds occurrences, and the relevant position of the sun during
each season of the year. Unfortunately, the lack of detailed data on cloud formation,
type and location is not allowing further analysis.

```{r relative-month-occurrences, echo=FALSE, fig.cap="Seasonal statistics of the number of CE occurrences for each month, normalized to the maximum occurrences on June. The box represents the values of the low $25\\,\\%$ percentile to $75\\,\\%$ percentile, where the thick horizontal line inside is the mean, the vertical lines extend to the maximum and minimum values, the dots are outlier values, and the rhombus is the mean."}
knitr::include_graphics("../images/clim-CE-month-norm-MAX-median-N-1.png")
```
<!--
```{r echo=FALSE, fig.cap="Distribution of the number of CE cases"}
knitr::include_graphics("../images/climCEmonth-1.png")
```
--->

@@

The distribution of the OIR intensity, follows an exponential decline
Figure\nobreakspace{}\ref{fig:ovir-distribution}, where there is an inverse relation
between the CE events frequency and their intensity. This is expected, as the stronger
the CE events become, the rarer the particular atmospheric and sun conditions can
occur. This fact, is indicative to the magnitude and the probability of the expected OIRs
events over Thessaloniki. Similar distribution of CE occurrences have been reported
by @Vamvakas2020, where the magnitude of OIR were
higher due to the location of the city of Patras, $2.5^\circ$ closer to the equator.

>>>>

Another interesting aspect of the CE events is their distribution within the year.
Figure\nobreakspace{}\ref{fig:relative-month-occurrences} shows the monthly box and
whisker plot of the CE number of occurrence normalized with the highest median value,
that occurs in June, depicting a clear seasonal cycle.  Although CE events are
present  throughout the year, the most active period  is during May and June.  During
the winter (December -- February), the number of CE cases is about $25\,\%$ of the
maximum, while in the intermediate months, the number of occurrences gradually ramps
between the maximum and minimum.  This seasonality is a combined effect of different
factors, among them the types of clouds, their frequency of occurrence the seasonally
varying relative position of the sun, as well as the local landscape characteristics
that may influence the formation of clouds.  Unfortunately, lack of detailed data on
cloud formation, type and location is not allowing further analysis. The interannual
variability of the monthly CE events is quite high as manifested by the large monthly
extremes, especially in the summer.

```{r relative-month-occurrences, echo=FALSE, fig.cap="Seasonal variability of the number of CE events normalized to the maximum occurring in June, in the form of a box and whisker plot. The monthly values have been also normalized to the relative abundance of valid GHI observations. The box contains the data between the lower $25\\,\\%$ and the upper $75\\,\\%$ percentiles,The thick horizontal line and the diamond symbol represent the median and the mean values, respectively. The vertical lines (whiskers) extend between the maximum and minimum monthly values and the solid circles are outliers."}
knitr::include_graphics("../images/clim-CE-month-norm-MAX-median-N-MW-1.png")
```
<!--
```{r echo=FALSE, fig.cap="Distribution of the number of CE cases"}
knitr::include_graphics("../images/climCEmonth-1.png")
```
--->

@@

The distribution of the number of CE events as a function of OIR is shown in
Figure\nobreakspace{}\ref{fig:ovir-distribution}. Apparently, there is an inverse
relation between the frequency of CE events  and OIR with an exponential-like
decline. This is expected, as the stronger the CE events become, the rarer are the
conditions favoring the occurrence of CE events. For the majority (over $60\%$) of
the CE events the OIR is below the long term average of $40\,\text{W}/\text{m}^2$,
while about $10\%$ of the events correspond to OIR larger than
$100\,\text{W}/\text{m}^2$ and up to highest value of 300 W/m^2^ .
particular atmospheric and sun conditions can
occur. This fact, is indicative to the magnitude and the probability of the expected OIRs
events over Thessaloniki. Similar distribution of CE occurrences have been reported
by @Vamvakas2020, where the magnitude of OIR were
higher due to the location of the city of Patras, $2.5^\circ$ closer to the equator.

### 2024-06-11 article/article.tex
Commit 2024-06-11 23:11

Another interesting aspect of the CE events is their distribution within
the year. Figure\nobreakspace{}\ref{fig:relative-month-occurrences}
shows the monthly box and whisker plot of the CE number of occurrence
normalized with the highest median value, that occurs on June. Although
CE events are present throughout the year, the most active period is
during May and June. During the winter (December -- February) the number
of CE cases is about \(25\,\%\) of the maximum. For the rest months, the
occurrences gradually ramp between the maximum and minimum. This
seasonality is a combinations of the clouds occurrences, and the
relevant position of the sun during each season of the year.
Unfortunately, the lack of detailed data on cloud formation, type and
location is not allowing further analysis.

@@

{\centering \includegraphics[width=1\linewidth]{../images/clim-CE-month-norm-MAX-median-N-1}

@@

\caption{Seasonal statistics of the number of CE occurrences for each month, normalized to the maximum occurrences on June. The box represents the values of the low $25\,\%$ percentile to $75\,\%$ percentile, where the thick horizontal line inside is the mean, the vertical lines extend to the maximum and minimum values, the dots are outlier values, and the rhombus is the mean.}\label{fig:relative-month-occurrences}
\end{figure}

The distribution of the OIR intensity, follows an exponential decline
Figure\nobreakspace{}\ref{fig:ovir-distribution}, where there is an
inverse relation between the CE events frequency and their intensity.
This is expected, as the stronger the CE events become, the rarer the
particular atmospheric and sun conditions can occur. This fact, is
indicative to the magnitude and the probability of the expected OIRs
events over Thessaloniki. Similar distribution of CE occurrences have
been reported by \citet{Vamvakas2020}, where the magnitude of OIR were
higher due to the location of the city of Patras, \(2.5^\circ\) closer
to the equator.

>>>>

Another interesting aspect of the CE events is their distribution within
the year. Figure\nobreakspace{}\ref{fig:relative-month-occurrences}
shows the monthly box and whisker plot of the CE number of occurrence
normalized with the highest median value, that occurs in June, depicting
a clear seasonal cycle. Although CE events are present throughout the
year, the most active period is during May and June. During the winter
(December -- February), the number of CE cases is about \(25\,\%\) of
the maximum, while in the intermediate months, the number of occurrences
gradually ramps between the maximum and minimum. This seasonality is a
combined effect of different factors, among them the types of clouds,
their frequency of occurrence the seasonally varying relative position
of the sun, as well as the local landscape characteristics that may
influence the formation of clouds. Unfortunately, lack of detailed data
on cloud formation, type and location is not allowing further analysis.
The interannual variability of the monthly CE events is quite high as
manifested by the large monthly extremes, especially in the summer.

@@

{\centering \includegraphics[width=1\linewidth]{../images/clim-CE-month-norm-MAX-median-N-MW-1}

@@

\caption{Seasonal variability of the number of CE events normalized to the maximum occurring in June, in the form of a box and whisker plot. The monthly values have been also normalized to the relative abundance of valid GHI observations. The box contains the data between the lower $25\,\%$ and the upper $75\,\%$ percentiles,The thick horizontal line and the diamond symbol represent the median and the mean values, respectively. The vertical lines (whiskers) extend between the maximum and minimum monthly values and the solid circles are outliers.}\label{fig:relative-month-occurrences}
\end{figure}

The distribution of the number of CE events as a function of OIR is
shown in Figure\nobreakspace{}\ref{fig:ovir-distribution}. Apparently,
there is an inverse relation between the frequency of CE events and OIR
with an exponential-like decline. This is expected, as the stronger the
CE events become, the rarer are the conditions favoring the occurrence
of CE events. For the majority (over \(60\%\)) of the CE events the OIR
is below the long term average of \(40\,\text{W}/\text{m}^2\), while
about \(10\%\) of the events correspond to OIR larger than
\(100\,\text{W}/\text{m}^2\) and up to highest value of 300
W/m\textsuperscript{2} . particular atmospheric and sun conditions can
occur. This fact, is indicative to the magnitude and the probability of
the expected OIRs events over Thessaloniki. Similar distribution of CE
occurrences have been reported by \citet{Vamvakas2020}, where the
magnitude of OIR were higher due to the location of the city of Patras,
\(2.5^\circ\) closer to the equator.
